### 2024-05-29 2024-5-13/sections/implementation.tex
add experiment results and explanation

We implemented the algorithm in Section~\ref{sec:algorithm} and Section~\ref{sec:heuristic} on top of OSTRICH, resulted to a string solver called $\ostrichrecl$. 
%
In this section, we evaluate the performance of $\ostrichrecl$ on three benchmark suites, that is, RegCoL, AutomatArk, and ComplexCounting. In the sequel, we first describe the three benchmark suites as well as the experiment setup. Then we present the experiment results. We evaluate the performance and correctness of our solver against the state-of-the-art string constraint solvers, including CVC5 \cite{cvc5}, Z3seq \cite{z3seq}, Z3str3 \cite{Z3-str3}, Z3str3RE \cite{BD+23}, and OSTRICH \cite{CHL+19}. We also compare $\ostrichrecl$ with two variants of it, that is, $\ostrichrecl_{\rm -ASR}$ and $\ostrichrecl_{\rm NUXMV}$, to justify the technical choices made in the decision procedure of Section~\ref{subsec:cefadec}.

@@

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. Other industrial benchmark suites are not utilized because they contain no counting operators. There are \denghang{todo} instances in total, and all benchmark instances are in the SMTLIB2 format.

@@

\medskip
\noindent
\emph{Distribution of problem instances w.r.t. counting bounds. }
The distribution of problem instances w.r.t. the counting bounds in RegCoL and AutomatArk suites is shown in Fig~\ref{fig:count_distri}, where the $x$-axis represents the counting bound and the $y$-axis represents the number of problem instances whose maximum counting bound is equal to the value of the $x$-axis. 
%The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). 
From Fig~\ref{fig:count_distri}, we can see that while most problem instances contain only small bounds, there are still around 3,000  (about 6\%) of them using large counting bounds (i.e. greater than or equal to $50$).
%Although most upper bounds are small, about two thousand of them are still greater than 50.

@@

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
\cite{Z3-str3}, Z3str3RE \cite{BD+23}, and OSTRICH
\cite{CHL+19}, on RegCoL, AutomatArk and ComplexCounting benchmark suites.
The experiment results can be found in Figure~\ref{fig:table_overall_eval}. Note that we take the results of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, and the results different from the ground truth are classified as \emph{soundness error}. We can see that $\ostrichrecl$ solves almost all \denghang{todo} instances, except 182 of them, that is, it solves 48,662 instances correctly. The number is 3,908/1,111/12,838/2,306/2,396 more than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (1.93s versus 1.62s).

@@

\subsection{Empirical justification of the technical choices and heuristics}\label{subsec:size_reduction_eval}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. Then we compare the performance of several heuristics in Section~\ref{sec:heuristic}.
\subsubsection{Empirical justification of the technical choices}
%We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
%Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
We compare $\ostrichrecl$ with $\ostrichrecl_{\rm -ASR}$ and $\ostrichrecl_{\rm NUXMV}$, where $\ostrichrecl_{\rm -ASR}$ is obtained from $\ostrichrecl$ by removing the automata size-reduction technique (i.e. Step 2 in Section~\ref{subsec:cefadec}), and $\ostrichrecl_{\rm NUXMV}$ is obtained from $\ostrichrecl$ by using the nuXmv model checker to solve the nonemptiness of $\cefadec$.

The experiment results can be found in Table~\ref{tab:results_simp}. 
We can see that $\ostrichrecl$ solves 1,503 more instances and is 2.21 times faster than $\ostrichrecl_{\rm -ASR}$. Therefore, the automata size-reduction technique indeed plays an essential role in the performance improvement. 
%To answer \textbf{Q2}, \textbf{the size-reduction techniques indeed play a vital role in the performance improvement}. 
Moreover, $\ostrichrecl$ solves 1,798 more instances and is 3.13 times faster than $\ostrichrecl_{\rm NUXMV}$. Therefore, the decision procedure in Section~\ref{subsec:cefadec} is more efficient to solve the $\cefadec$ problem than nuXmv. 
\begin{table}[H]
  \import{tables}{table_simp.tex}
  \caption{Empirical justification of the technical choices in the decision procedure}\label{tab:results_simp}
\end{table}

\subsubsection{Empirical justification of the heuristics} We compare the performance of $\ostrichrecl$ with ${\rm -NEST}$, ${\rm -COMP}$, ${\rm -FIND}$, and ${\rm -ALL}$ where ${\rm -NEST}$ is obtained from $\ostrichrecl$ by removing the heuristic for nested counting, ${\rm -COMP}$ is obtained from $\ostrichrecl$ by removing the heuristic for complement on counting operators, ${\rm -FIND}$ is obtained from $\ostrichrecl$ by removing the heuristic for finding the accepted word of the automaton, and ${\rm -ALL}$ is obtained from $\ostrichrecl$ by removing all the heuristics.
The experiment results can be found in Table~\ref{tab:results_heuristic}.

\begin{table}
  \import{tables}{table_heuristics.tex}
  \caption{Empirical justification of the heuristics}\label{tab:results_heuristic}
\end{table}

>>>>

We implemented the algorithm in Section~\ref{sec:algorithm} and Section~\ref{sec:heuristic} on top of OSTRICH, resulted to a string solver called $\ostrichrecl$. 
%
In this section, we evaluate the performance of $\ostrichrecl$ on three benchmark suites, that is, RegCoL, AutomatArk, and ComplexCounting. In the sequel, we first describe the three benchmark suites as well as the experiment setup. Then we present the experiment results. We evaluate the performance and correctness of our solver against the state-of-the-art string constraint solvers, including CVC5 \cite{cvc5}, Z3seq \cite{z3seq}, Z3str3 \cite{Z3-str3}, Z3str3RE \cite{BD+23}, and OSTRICH \cite{CHL+19}. We also compare $\ostrichrecl$ with several variants of it, to justify the technical choices and heuristics we take in the Section~\ref{subsec:cefadec} and Section~\ref{sec:heuristic}.

@@

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. Other industrial benchmark suites are not utilized because they contain no counting operators. There are 49,843 instances in total, where 48,843 of them are collected from the real world and 1,000 of them are generated by fuzzing tool. All benchmark instances are in the SMTLIB2 format.

@@

\medskip
\noindent
\emph{Distribution of problem instances w.r.t. counting bounds. }
The distribution of problem instances w.r.t. the counting bounds in the three suites is shown in Fig~\ref{fig:count_distri}, where the $x$-axis represents the counting bound and the $y$-axis represents the number of problem instances containing the counting bound. 
%The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). 
From Fig~\ref{fig:count_distri}, we can see that while most problem instances contain only small bounds, there are still around 3,000  (about 6\%) of them using large counting bounds (i.e. greater than or equal to $50$).
%Although most upper bounds are small, about two thousand of them are still greater than 50.

@@

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
\cite{Z3-str3}, Z3str3RE \cite{BD+23}, and OSTRICH
\cite{CHL+19}, on RegCoL, AutomatArk and ComplexCounting benchmark suites.
The experiment results can be found in Figure~\ref{fig:table_overall_eval}. Note that we take the results of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, and the results different from the ground truth are classified as \emph{error}. We can see that $\ostrichrecl$ solves almost all instances, except 182 of them, that is, it solves 49,638 instances correctly. The number is 4,547/1,917/13,618/2,764/3,134 more than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (2.88s versus 1.99s).

@@

\subsection{Empirical justification of the technical choices and heuristics}\label{subsec:size_reduction_eval}
We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 of Section~\ref{subsec:cefadec}. We practically answer why we choose LIA formula to solve the nonemptiness problem instead of using the nuXmv model checker like \cite{atva2020}. Finally, we carry out experiment on the ComplexCounting benchmarks to justify the heuristics of Section~\ref{sec:heuristic}.
\subsubsection{Empirical justification of the technical choices}
%We also do experiments to evaluate the effectiveness of the automata size-reduction techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
%Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the size-reduction techniques removed. 
We compare $\ostrichrecl$ with its variants ${\rm -ASR}$ and ${\rm +NUXMV}$. ${\rm -ASR}$ is obtained from $\ostrichrecl$ by removing the automata size-reduction technique (i.e. Step 2 of Section~\ref{subsec:cefadec}). ${\rm +NUXMV}$ uses the nuXmv model checker to solve the nonemptiness of $\cefadec$.

The experiment results can be found in Table~\ref{tab:results_heuristic}.
We can see that $\ostrichrecl$ solves 1,083 more instances and is 1.65 times faster than ${\rm -ASR}$. Therefore,  the automata size-reduction technique indeed plays an essential role in the performance improvement.
We can see that $\ostrichrecl$ solves 2,532 more instances and is 2.38 times faster than ${\rm +NUXMV}$. Therefore, the decision procedure in Section~\ref{subsec:cefadec} is more efficient to solve the $\cefadec$ problem than nuXmv. 
\begin{table}
  \import{tables}{table_heuristics.tex}
  \caption{Empirical justification of the technical choices and heuristics}\label{tab:results_heuristic}
\end{table}

\subsubsection{Empirical justification of the heuristics} 
We also compare the performance of $\ostrichrecl$ with its variants ${\rm -NEST}$, ${\rm -COMP}$, ${\rm -FIND}$. ${\rm -NEST}$ is obtained from $\ostrichrecl$ by removing the heuristic for nested counting. ${\rm -COMP}$ is obtained by removing the heuristic for complement on counting operators. ${\rm -FIND}$ is obtained by removing the heuristic for finding the accepted word of the automaton.

The overall experiment results can be found in Table~\ref{tab:results_heuristic}. We can see that $\ostrichrecl$ solves 77/388/349 more instances and is 1.13/1.18/1.24 times faster than ${\rm -NEST}$/${\rm -COMP}$/${\rm -FIND}$ respectively.

The ComplexCounting benchmark experiment results is shown in Table~\ref{tab:results_heuristic_complex}. We can see that $\ostrichrecl$ solves 82/339/326 more instances and is 1.02/1.92/1.61 times faster than ${\rm -NEST}$/${\rm -COMP}$/${\rm -FIND}$ respectively.

From the results, the heuristic for complement on counting operators and the heuristic for finding the accepted word of the automaton are more effective than the heuristic for nested counting. Futhermore, the heuristic for complement on counting operators contributes both to satisfiable and unsatisfiable instances, while the heuristic for finding the accepted word and the heuristic for nested counting contribute only to satisfiable instances.
All the heuristics improve the performance of $\ostrichrecl$.

\begin{table}
  \import{tables}{table_heuristics_complex.tex}
  \caption{Empirical justification of the heuristics on ComplexCounting benchmarks}\label{tab:results_heuristic_complex}
\end{table}
